Detection of Oscillations in Aldebaran with Ground-Based Observations

### 2017-09-20 paper/aldebaran.tex
intro

Aldebaran, or $\alpha$~Tauri, is one of the brightest and most easily recognized naked-eye stars, and has long been the subject of astronomical investigations. Following the proposal of \citet{struverv} several groups began to search for planets by the radial velocity (RV) method, looking for Doppler shifts from the star's reflex motion around the common centre of mass, culminating in the discovery of 51~Peg~b \citep{51peg}, the first exoplanet to be recognized as such. Before this, \citet{hatzes1993} noted RV variations in Pollux \citep[$\beta$~Gem; subsequently confirmed as a planet:][]{betgemconf,betgemconf2}, Arcturus ($\alpha$~Boo, unconfirmed), and Aldebaran.

@@

\section{Aldebaran}
Aldebaran is one of the nearest red giant stars, at a distance of only $19.96 \pm 0.38$~pc \citep{hipparcos}, and its position near the Ecliptic permits the determination of its angular diameter by lunar occultations and by interferometry \citep[$20.58 \pm 0.03$ mas;][]{richichi2005,1979ApJ...228L.111B,brown1979,panek1980}.

>>>>

Aldebaran, or $\alpha$~Tauri, is a well-known first-magnitude naked-eye red giant star, and has long been the subject of astronomical investigations. Following the proposal of \citet{struverv} several groups began to search for planets by the radial velocity (RV) method, looking for Doppler shifts from the star's reflex motion around the common centre of mass, culminating in the discovery of 51~Peg~b \citep{51peg}, the first exoplanet to be recognized as such. Before this, \citet{hatzes1993} noted RV variations in Pollux \citep[$\beta$~Gem; subsequently confirmed as a planet:][]{betgemconf,betgemconf2}, Arcturus ($\alpha$~Boo, unconfirmed), and Aldebaran. After further investigation by \citet{Hatzes1998}, \citet{Hatzes2015} now claim a firm RV detection of a planetary-mass companion Aldebaran~b, with a period of $628.96 \pm 0.90$~d.

In this paper, we present a re-analysis of these original RV data in which we not only confirm this signal, but detect acoustic oscillations in Aldebaran for the first time. We validate this method and its result with new RV observations with the SONG~Telescope, and photometry from the \ktwo Mission. By measuring the average frequency \numax of these \emph{p}-mode oscillations we asteroseismically determine the mass of Aldebaran to be $1.17 \pm 0.05$~\msun. This measured stellar mass allows us to calculate that any moons of this giant planet, although they are now likely to be very hot, may have had equilibrium temperatures comparable to that of the Earth when Aldebaran was on the main sequence, raising the possibility that tehy may have once been habitable.

Our new approach to asteroseismic data analysis, based on Continuous Auto-Regressive Moving Average (CARMA) models, can extract exoplanet signals together with measures of \numax from sparse and irregularly-sampled time series. An all-sky survey planetary companions to precisely measure the masses of all nearby red giant stars is feasible with this new approach, and the required data either already exist in large exoplanet surveys, or are easy to obtain with ground-based telescopes.

@@

\section{Aldebaran}
Aldebaran is a red giant star with spectral type K5, one of the nearest such stars at a distance of only $19.96 \pm 0.38$~pc \citep{hipparcos}. Its position near the Ecliptic permits the determination of its angular diameter by lunar occultations and by interferometry \citep[$20.58 \pm 0.03$ mas;][]{richichi2005,1979ApJ...228L.111B,brown1979,panek1980}.
